---
title: "Getting started with Logger"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with Logger}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

## Introduction

The `Logger` package provides a flexible and powerful logging system for R applications. It includes a `Logger` class for creating customisable loggers, as well as helper functions for debugging and error reporting. This vignette will guide you through the basics of using the `Logger` package and demonstrate how to leverage its features to improve your R workflows.

## Installation

You can install the `Logger` package from CRAN with:

```{r, eval = FALSE}
remotes::install_github("Logger")
```

## Basic Usage

First, let's load the package and create a basic logger:

```{r}
box::use(Logger[Logger, LogLevel])

# Create a basic logger
log <- Logger$new()

# Log some messages
log$info("This is an informational message")
log$warn("This is a warning")
log$error("This is an error")
```

## Customising the Logger

You can customise the logger by specifying the minimum log level, output file, and custom print function:

```{r}
# Create a custom logger
custom_log <- Logger$new(
    level = LogLevel$WARNING,
    file_path = "app.log",
    print_fn = message
)

custom_log$info("This won't be logged")
custom_log$warn("This will be logged to console and file")
custom_log$error("This is an error message")
```

## Using Helper Functions

The `Logger` package includes several helper functions that can be used in conjunction with the `Logger` class to provide more detailed information in your logs. Let's explore how to use these functions effectively.

### Finding and Logging Data Issues

Suppose we have a dataset with some problematic values, and we want to log where these issues occur. We can use the `valueCoordinates` function to locate the problematic values and include this information in our log messages.

```{r}
box::use(Logger[valueCoordinates])

# Create a sample dataset with some issues
df <- data.frame(
    a = c(1, NA, 3, 4, 5),
    b = c(2, 4, NA, 8, 10),
    c = c(3, 6, 9, NA, 15)
)

# Create a logger
log <- Logger$new()

# Find coordinates of NA values
na_coords <- valueCoordinates(df)

if (nrow(na_coords) > 0) {
    log$warn(
        "NA values found in the dataset",
        data = list(
            na_locations = na_coords
        )
    )
}
```

This will produce a log entry like:

### Logging Errors with Context

When an error occurs, it's often useful to log not just the error message, but also the context in which the error occurred. Here's an example of how to do this using the `Logger` class and helper functions:

```{r}
box::use(Logger[tableToString])

log <- Logger$new()

process_data <- function(df) {
    tryCatch({
        result <- df$a / df$b
        if (any(is.infinite(result))) {
            inf_coords <- valueCoordinates(data.frame(result), Inf)
            log$error(
                "Division by zero occurred",
                data = list(
                    infinite_values = inf_coords,
                    dataset_preview = tableToString(df)
                )
            )
            stop("Division by zero error")
        }
        return(result)
    }, error = function(e) {
        log$error(
            paste("An error occurred while processing data:", e$message),
            data = list(dataset_preview = tableToString(df)),
            error = e
        )
        stop(e)
    })
}

# Test the function with problematic data
df <- data.frame(a = c(1, 2, 3), b = c(0, 2, 0))
process_data(df)
```

## Logging in Parallel Environments

When working with parallel processing, standard logging functions might not work as expected. The `Logger` package provides a `messageParallel` function to ensure messages are properly logged from parallel processes:

```{r}
box::use(future)
box::use(future.apply[future_lapply])

log <- Logger$new(print_fn = messageParallel)

future::plan(future::multisession, workers = 2)

result <- future_lapply(1:5, function(i) {
    messageParallel(sprintf("Processing item %d", i))
    if (i == 3) {
        log$warn(sprintf("Warning for item %d", i))
    }
    return(i * 2)
})

future::plan(future::sequential)
```

This ensures that messages from parallel processes are properly captured and logged.

## Conclusion

The `Logger` package provides a robust and flexible logging system for R applications. By using the `Logger` class in combination with helper functions like `valueCoordinates` and `tableToString`, you can create informative and context-rich log messages that greatly aid in debugging and monitoring your R scripts and applications.

Remember to adjust the log level, output file, and other settings to suit your specific needs. Happy logging!